Appendix0_10

### 2018-08-17 paper/Appendices/Appendix0_10.Rmd
appendix C figlevel 1 CW4

```{r eval=FALSE}
# NDNS analysis, data management ------------------------------------------

# Change the data path accordingly ----------------------------------------
setwd("/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/") # on Ubuntu
library(epiDisplay)
library(plyr)
library(tidyverse)

# Read the data into memory -----------------------------------------------
library(haven)
data <- read_dta("ndns_rp_yr1-4a_foodleveldietarydata_uk.dta")
data56 <- read_dta("ndns_rp_yr5-6a_foodleveldietarydata.dta")
data78 <- read_dta("ndns_rp_yr7-8a_foodleveldietarydata.dta")
names(data)
names(data56)
names(data78)
names(data)[names(data) == "seriali"] <- "id"
names(data56)[names(data56) == "seriali"] <- "id"
names(data78)[names(data78) == "seriali"] <- "id"

# Extract the data we needed ----------------------------------------------
df14d <- data[, c(113, 1, 2, 3, 5, 6, 7, 8, 9, 21, 24, 55, 57, 58, 
    59, 60, 61, 62, 63, 64)]
var <- names(df14d)
df56d <- data56 %>% select(var)
df78d <- data78 %>% select(var)
dfs1 <- rbind(df14d, df56d, df78d)
dfs2 <- dfs1[dfs1$Age >= 19, ]
rm(data, data56, data78)
dfs2

@@

dfs2$MealTime_chr <- as.character(dfs2$MealTime)
dfs2$MealTime_hm <- unlist(strsplit(dfs2$MealTime_chr, " "))[c(FALSE, 
    TRUE)]
dfs2$MealHourN <- as.numeric(unlist(strsplit(dfs2$MealTime_hm, ":"))[c(TRUE, 
    FALSE, FALSE)])
dfs2$MealMinN <- as.numeric(unlist(strsplit(dfs2$MealTime_hm, ":"))[c(FALSE, 
    TRUE, FALSE)])
dfs2$MealMinN0 <- (60 * dfs2$MealHourN) + dfs2$MealMinN
dfs3 <- dfs2[order(dfs2$id, dfs2$DayNo, dfs2$MealMinN0), ]
length(unique(dfs3$id))  ## number of participants = 6155

# Create a subset data with only the first observation of each
# participant --------
NDNS <- dfs3[!duplicated(dfs3$id), ]
with(NDNS, tab1(SurveyYear, graph = FALSE, decimal = 2))

# #SurveyYear :
            # Frequency Percent Cum. percent
# NDNS Year 1       801   13.01        13.01
# NDNS Year 2       812   13.19        26.21
# NDNS Year 3       782   12.71        38.91
# NDNS Year 4      1055   17.14        56.05
# NDNS Year 5       625   10.15        66.21
# NDNS Year 6       663   10.77        76.98
# NDNS Year 7       703   11.42        88.40
# NDNS Year 8       714   11.60       100.00
  # Total          6155  100.00       100.00

# create a variable combine id and day No ---------------------------------
dfs3 <- dfs3 %>%
mutate(id_dy = paste(id, DayNo, sep = "D"))

# For each subject, the total energy/carbohydrate intake for each eating 
# time can be calculated --------
old <- Sys.time()
Energy <- ddply(dfs3, .(id_dy, id, SurveyYear, DayNo, Age, Sex, 
                        DiaryDaysCompleted, MealHourN, DayofWeek), 
                summarise, 
                Tot_Energ = sum(EnergykJ), 
                Tot_Carb = sum(Carbohydrateg), 
                Tot_Sugar = sum(Totalsugarsg), 
                Tot_Starch = sum(Starchg))
new <- Sys.time() - old
print(new)
# Time difference of 3.876385 mins

rm(df14d, df56d, df78d, dfs2)

# Calculate the energy from total carbohydrates ---------------------------
Energy <- Energy %>% 
  mutate(KJcarbo = Tot_Carb * 16) %>% 
  mutate(CarKJpercentage = KJcarbo/Tot_Energ) %>% 
  mutate(Carbo = cut(CarKJpercentage, breaks = c(0, 0.26, 0.75, 2), 
        right = FALSE)) %>% mutate(Carbo2 = cut(CarKJpercentage, breaks = c(0, 
    0.26, 2), right = FALSE))
Energy0 <- Energy[!(Energy$Tot_Energ == 0), ]  
          # some food consumption does not contain any carbohydrates
Energy0$Carbo <- factor(Energy0$Carbo, labels = c("Low_carb", "Med_carb", 
    "High_carb"))
Energy0$Carbo2 <- factor(Energy0$Carbo2, labels = c("Low_carb", "Med_or_high_carb"))

# Generate data sets for each day ----------------------------------------------------------
dta_day1 <- Energy0 %>% 
  filter(DayNo == 1) %>% 
  select(c("id", "Age", 
    "Sex", "DayofWeek", "MealHourN", "Carbo", "Carbo2")) %>% 
  mutate(DayofWeek = factor(DayofWeek, 
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", 
        "Saturday", "Sunday")))

dta_day2 <- Energy0 %>% 
  filter(DayNo == 2) %>% 
  select(c("id", "Age", 
    "Sex", "DayofWeek", "MealHourN", "Carbo", "Carbo2")) %>% 
  mutate(DayofWeek = factor(DayofWeek, 
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", 
        "Saturday", "Sunday")))

dta_day3 <- Energy0 %>% 
  filter(DayNo == 3) %>% 
  select(c("id", "Age", 
    "Sex", "DayofWeek", "MealHourN", "Carbo", "Carbo2")) %>% 
  mutate(DayofWeek = factor(DayofWeek, 
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", 
        "Saturday", "Sunday")))

dta_day4 <- Energy0 %>% 
  filter(DayNo == 4) %>% 
  select(c("id", "Age", 
    "Sex", "DayofWeek", "MealHourN", "Carbo", "Carbo2")) %>% 
  mutate(DayofWeek = factor(DayofWeek, 
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", 
        "Saturday", "Sunday")))
vecid1 <- unique(dta_day1$id) # n = 6153
vecid2 <- unique(dta_day2$id) # n = 6153
vecid3 <- unique(dta_day3$id) # n = 6151
vecid4 <- unique(dta_day4$id) # n = 6026

Noday1 <- setdiff(vecid, vecid1)  # two subjects did not have day 1 data
Noday2 <- setdiff(vecid, vecid2)  # two subjects did not have day 2 data
Noday3 <- setdiff(vecid, vecid3)  # four subjects did not have day 3 data
Noday4 <- setdiff(vecid, vecid4)  # 129 subjects did not have day 4 data

# Transform the data shape from long to wide ---------------------------------
dta_d1_wide <- dta_day1[, -7] %>% 
  spread(key = MealHourN, value = Carbo)
names(dta_d1_wide)[5:28] <- paste(rep("H", 24), 0:23, sep = "")

dta_d2_wide <- dta_day2[, -7] %>% 
  spread(key = MealHourN, value = Carbo)
names(dta_d2_wide)[5:28] <- paste(rep("H", 24), 0:23, sep = "")

dta_d3_wide <- dta_day3[, -7] %>% 
  spread(key = MealHourN, value = Carbo)
names(dta_d3_wide)[5:28] <- paste(rep("H", 24), 0:23, sep = "")

dta_d4_wide <- dta_day4[, -7] %>% 
  spread(key = MealHourN, value = Carbo)
names(dta_d4_wide)[5:28] <- paste(rep("H", 24), 0:23, sep = "")

# recode NA to not eating -------------------------------------------------
for (i in 5:ncol(dta_d1_wide)) 
  if (is.factor(dta_d1_wide[, i])) levels(dta_d1_wide[, 
    i]) <- c(levels(dta_d1_wide[, i]), "Not_eating")

dta_d1_wide[is.na(dta_d1_wide)] <- "Not_eating"

for (i in 5:ncol(dta_d2_wide)) 
  if (is.factor(dta_d2_wide[, i])) levels(dta_d2_wide[, 
    i]) <- c(levels(dta_d2_wide[, i]), "Not_eating")

dta_d2_wide[is.na(dta_d2_wide)] <- "Not_eating"

for (i in 5:ncol(dta_d3_wide)) 
  if (is.factor(dta_d3_wide[, i])) levels(dta_d3_wide[, 
    i]) <- c(levels(dta_d3_wide[, i]), "Not_eating")

dta_d3_wide[is.na(dta_d3_wide)] <- "Not_eating"

for (i in 5:ncol(dta_d4_wide)) 
  if (is.factor(dta_d4_wide[, i])) levels(dta_d4_wide[, 
    i]) <- c(levels(dta_d4_wide[, i]), "Not_eating")

dta_d4_wide[is.na(dta_d4_wide)] <- "Not_eating"
```

@@

>>>>

```{r eval=FALSE}

# NDNS analysis, data management ------------------------------------------

# Change the data path accordingly ----------------------------------------

setwd(".../UKDA-6533-stata11_se/stata11_se/") 
setwd("~/Downloads/UKDA-6533-stata11_se/stata11_se/")

library(epiDisplay)
library(plyr)
library(tidyverse)

# Read the data into memory -----------------------------------------------

library(haven)
data <- read_dta("ndns_rp_yr1-4a_foodleveldietarydata_uk.dta")
data56 <- read_dta("ndns_rp_yr5-6a_foodleveldietarydata.dta")
data78 <- read_dta("ndns_rp_yr7-8a_foodleveldietarydata.dta")

names(data)
names(data56)
names(data78)
names(data)[names(data)=="seriali"] <- "id"
names(data56)[names(data56)=="seriali"] <- "id"
names(data78)[names(data78)=="seriali"] <- "id"

# Extract the variables needed -------------------------------------------

df14d <- data[,c(113,1,2,3,5,6,7,8,9,21, 22, 23, 24, 53, 55,
                 57,58,59,60,61,62,63,64,65)]
var <- names(df14d)
df56d <- data56 %>% 
  select(var)
df78d <- data78 %>% 
  select(var)
dfs1 <- rbind(df14d, df56d, df78d)

dfs2 <- dfs1[dfs1$Age>=19,] # keep participants who aged 19 or older
rm(data, data56, data78) # remove the unneeded big dataset

dfs2

@@

dfs2$MealTime_chr <- as.character(dfs2$MealTime)
dfs2$MealTime_hm <- unlist(strsplit(dfs2$MealTime_chr," "))[c(FALSE, 
                                                              TRUE)]
dfs2$MealHourN <- as.numeric(unlist(strsplit(dfs2$MealTime_hm,
                                             ":"))[c(TRUE, FALSE, FALSE)])
dfs2$MealMinN <- as.numeric(unlist(strsplit(dfs2$MealTime_hm,
                                            ":"))[c(FALSE, TRUE, FALSE)])

dfs2$MealMinN0 <- (60*dfs2$MealHourN)+dfs2$MealMinN

dfs3 <- dfs2[order(dfs2$id,dfs2$DayNo,dfs2$MealMinN0),]

length(unique(dfs3$id)) ## number of participants = 6155

# Create a subset data with only the first observation of each participant --------

NDNS <- dfs3[!duplicated(dfs3$id), ]
with(NDNS, tab1(SurveyYear, graph = FALSE, decimal = 2))
```

```
# #SurveyYear : 
#             Frequency Percent Cum. percent
# NDNS Year 1       801   13.01        13.01
# NDNS Year 2       812   13.19        26.21
# NDNS Year 3       782   12.71        38.91
# NDNS Year 4      1055   17.14        56.05
# NDNS Year 5       625   10.15        66.21
# NDNS Year 6       663   10.77        76.98
# NDNS Year 7       703   11.42        88.40
# NDNS Year 8       714   11.60       100.00
#   Total          6155  100.00       100.00
```

```{r eval=FALSE}
# how many men and women --------------------------------------------------

with(NDNS, tab1(Sex, graph = FALSE, decimal = 2))
```

```
# Sex : 
#          Frequency Percent Cum. percent
# 1            2537   41.22        41.22 Men
# 2            3618   58.78       100.00 Women
#   Total      6155  100.00       100.00
```

```{r eval=FALSE}
# create a variable combine id and day No ---------------------------------

dfs3 <- dfs3 %>% 
  mutate(id_dy = paste(id, DayNo, sep = "D"))

# For each subject, the total energy/carbohydrate intake for each 
# eating time can be calculated --------

old<-Sys.time()
Energy <- ddply(dfs3, .(id_dy, id, SurveyYear, DayNo, Age, Sex, 
                        DiaryDaysCompleted, MealHourN, DayofWeek),  
                summarise, Tot_Energ = sum(EnergykJ), 
                Tot_Carb = sum(Carbohydrateg), 
                Tot_Sugar = sum(Totalsugarsg), 
                Tot_Starch = sum(Starchg), 
                Tot_Fibre = sum(Englystfibreg),
                Tot_Fat   = sum(Fatg), 
                Tot_Prot  = sum(Proteing), 
                Tot_Alc   = sum(Alcoholg), 
                Tot_NMES  = sum(Nonmilkextrinsicsugarsg))
new<-Sys.time()-old
print(new)
# Time difference of 6.429822 min

# reset the time intervals into time slots --------------------------------

### Breakfast:       6am to 9am
### morning snack:   9am to 12noon
### lunch:           12noon to 2pm
### afternoon snack: 2pm to 5pm
### dinner:          5pm to 8pm
### night snack:     8pm to 10pm
### midnight:        10pm to 6am

Energy <- Energy %>% 
  mutate(TimeSlot = cut(MealHourN, breaks = c(6, 9, 12, 14, 17, 20, 22), 
                        right = FALSE))

levels(Energy$TimeSlot) <- c(levels(Energy$TimeSlot), "[22, 6)")

Energy$TimeSlot[is.na(Energy$TimeSlot)] <- "[22, 6)"

tab1(Energy$TimeSlot)

# For each subject, the total energy/carbohydrate intake for each 
# time slot can be calculated --------
old<-Sys.time()
Energy <- ddply(Energy, .(id_dy, id, SurveyYear, DayNo, Age, Sex, 
                          DiaryDaysCompleted, TimeSlot, DayofWeek),  
                summarise, 
                Tot_Energ = sum(Tot_Energ), 
                Tot_Carb = sum(Tot_Carb), 
                Tot_Sugar = sum(Tot_Sugar),
                Tot_Starch = sum(Tot_Starch), 
                Tot_Fibre = sum(Tot_Fibre), 
                Tot_Fat   = sum(Tot_Fat), 
                Tot_Prot  = sum(Tot_Prot), 
                Tot_Alc   = sum(Tot_Alc), 
                Tot_NMES  = sum(Tot_NMES))
new<-Sys.time()-old
print(new)
# Time difference of 3.74195 mins

# Calculate the energy from total carbohydrates ---------------------------

Energy <- Energy %>% 
  mutate(KJcarbo = Tot_Carb*16) %>% 
  mutate(CarKJpercentage = KJcarbo/Tot_Energ) %>% 
  mutate(Carbo = cut(CarKJpercentage, breaks = c(0, 0.50, 2), right = FALSE))

Energy0 <- Energy[!(Energy$Tot_Energ == 0), ] # discard those eating occassion 
# with 0 energy intake 
write.csv(Energy0, file = "Energy_slots.csv") # for later analysis

Energy0$Carbo <- factor(Energy0$Carbo, labels = c("< 50%", ">= 50%"))

vecid<-unique(Energy0$id)

# Filter the data by observation day--------------------------------------------

dta_day1 <- Energy0 %>% 
  filter(DayNo == 1) %>% 
  select(c("id", "id_dy", "Age", "Sex", 
           "DayofWeek",  "TimeSlot", "Carbo")) %>%  
  mutate(DayofWeek = factor(DayofWeek, 
                            levels = c("Monday", "Tuesday", 
                                       "Wednesday", "Thursday", 
                                       "Friday", "Saturday", "Sunday")))

dta_day2 <- Energy0 %>% 
  filter(DayNo == 2) %>% 
  select(c("id", "id_dy", "Age", "Sex", 
           "DayofWeek",  "TimeSlot", "Carbo")) %>%  
  mutate(DayofWeek = factor(DayofWeek, 
                            levels = c("Monday", "Tuesday", 
                                       "Wednesday", "Thursday", 
                                       "Friday", "Saturday", "Sunday")))

dta_day3 <- Energy0 %>% 
  filter(DayNo == 3) %>% 
  select(c("id", "id_dy", "Age", "Sex", 
           "DayofWeek",  "TimeSlot", "Carbo")) %>%  
  mutate(DayofWeek = factor(DayofWeek, 
                            levels = c("Monday", "Tuesday", 
                                       "Wednesday", "Thursday", 
                                       "Friday", "Saturday", "Sunday")))

dta_day4 <- Energy0 %>% 
  filter(DayNo == 4) %>% 
  select(c("id", "id_dy", "Age", "Sex", 
           "DayofWeek",  "TimeSlot", "Carbo")) %>%  
  mutate(DayofWeek = factor(DayofWeek, 
                            levels = c("Monday", "Tuesday", 
                                       "Wednesday", "Thursday", 
                                       "Friday", "Saturday", "Sunday")))

vecid1<-unique(dta_day1$id) # n = 6153
vecid2<-unique(dta_day2$id) # n = 6153
vecid3<-unique(dta_day3$id) # n = 6151
vecid4<-unique(dta_day4$id) # n = 6026

Noday1 <- setdiff(vecid, vecid1) # two subjects did not have day 1 data

Noday2 <- setdiff(vecid, vecid2) # two subjects did not have day 2 data

Noday3 <- setdiff(vecid, vecid3) # four subjects did not have day 3 data

Noday4 <- setdiff(vecid, vecid4) # 129 subjects did not have day 4 data

# Long to wide data -------------------------------------------------------

dta_d1_wide <- dta_day1 %>% 
  spread(key = TimeSlot, 
         value = Carbo)

head(dta_d1_wide)
names(dta_d1_wide)[6:12] <- c("H6_9", "H9_12", "H12_14", "H14_17", 
                              "H17_20", "H20_22", "H22_6")
names(dta_d1_wide)

dta_d2_wide <- dta_day2 %>% 
  spread(key = TimeSlot, 
         value = Carbo)

head(dta_d2_wide)
names(dta_d2_wide)[6:12] <- c("H6_9", "H9_12", "H12_14", "H14_17", 
                              "H17_20", "H20_22", "H22_6")
names(dta_d2_wide)

dta_d3_wide <- dta_day3 %>% 
  spread(key = TimeSlot, 
         value = Carbo)

head(dta_d3_wide)
names(dta_d3_wide)[6:12] <- c("H6_9", "H9_12", "H12_14", "H14_17", 
                              "H17_20", "H20_22", "H22_6")
names(dta_d3_wide)

dta_d4_wide <- dta_day4 %>% 
  spread(key = TimeSlot, 
         value = Carbo)

head(dta_d4_wide)
names(dta_d4_wide)[6:12] <- c("H6_9", "H9_12", "H12_14", "H14_17", 
                              "H17_20", "H20_22", "H22_6")
names(dta_d4_wide)

# recode NA to not eating -------------------------------------------------

for (i in 6:ncol(dta_d1_wide)) 
  if(is.factor(dta_d1_wide[,i])) 
    levels(dta_d1_wide[,i]) <- c("1", "2", "0")

dta_d1_wide[is.na(dta_d1_wide)] <- "0"

for (i in 6:ncol(dta_d2_wide)) 
  if(is.factor(dta_d2_wide[,i])) 
    levels(dta_d2_wide[,i]) <- c("1", "2",  "0")

dta_d2_wide[is.na(dta_d2_wide)] <- "0"

for (i in 6:ncol(dta_d3_wide)) 
  if(is.factor(dta_d3_wide[,i])) 
    levels(dta_d3_wide[,i]) <-  c("1", "2", "0")

dta_d3_wide[is.na(dta_d3_wide)] <- "0"

for (i in 6:ncol(dta_d4_wide)) 
  if(is.factor(dta_d4_wide[,i])) 
    levels(dta_d4_wide[,i]) <-  c("1", "2", "0")

dta_d4_wide[is.na(dta_d4_wide)] <- "0"

dta_all <- rbind(dta_d1_wide, dta_d2_wide, dta_d3_wide, dta_d4_wide)

dta_all <- dta_all[order(dta_all$id,dta_all$id_dy),]

# Export the data for Mplus -------------------------------------------------

write_csv(dta_all, path = "dta_NDNS_Tslots.csv")
write_delim(dta_all, "dta_NDNS_Tslots.dat", na = ".", delim = " ")
```

@@

```{r  eval=FALSE}
library(plyr)
library(epiDisplay)
library(tidyverse)
library(dplyr)
library(readr)
library(haven)
library(naniar)

# read the individual level data sets ----------------------------------------

blood78 <- read_dta("ndns_rp_yr7-8a_indiv.dta")
blood56 <- read_dta("ndns_rp_yr5-6a_indiv.dta")
blood14 <- read_dta("ndns_rp_yr1-4a_indiv_uk.dta")

food14 <- read_dta("ndns_rp_yr1-4a_personleveldietarydata_uk.dta")

food56 <- read_dta("ndns_rp_yr5-6a_personleveldietarydata.dta")

food78 <- read_dta("ndns_rp_yr7-8a_personleveldietarydata.dta")

names(blood78)[names(blood78)=="seriali"] <- "ID"
names(blood56)[names(blood56)=="seriali"] <- "ID"
names(blood14)[names(blood14)=="seriali"] <- "ID"
names(food78)[names(food78)=="seriali"] <- "ID"
names(food56)[names(food56)=="seriali"] <- "ID"
names(food14)[names(food14)=="seriali"] <- "ID"

# Loading the data from Mplus output
CW3CB3 <- read_table2("../NDNSslot_CW3CB3.txt",  # change the path accordingly
                      col_names = FALSE)

names(CW3CB3) <- c("Breakfast", "Morning.snack", "Lunch",
                   "Afternoon.snack", "Dinner", "Before.bedtime.snack",
                   "Midnight.food", "ID_DAY", "AGE", "SEX", "CPROB1",
                   "CPROB2", "CPROB3", "CPROB4", "CPROB5", "CPROB6", 
                   "CPROB7", "CPROB8", "CPROB9", 
                   "CB", "CW", "MLCJOINT", "ID")

CW3idday <- CW3CB3 %>% 
  select(ID, ID_DAY, CW, CB, MLCJOINT, AGE, SEX)

Energy_slots <- read_csv("../Energy_slots.csv") # change the path accordingly
Energy_slots <- Energy_slots %>% 
  rename(ID = id)

# Recode day level classification to keep consistency
CW3idday$CW_new <- 0
CW3idday$CW_new[CW3idday$CW == 1] <- 3
CW3idday$CW_new[CW3idday$CW == 2] <- 1
CW3idday$CW_new[CW3idday$CW == 3] <- 2

Energy_slots <- Energy_slots %>% 
  left_join(CW3idday, by = c("ID", "DayofWeek"))

Energy_slots$TimeSlot <- factor(Energy_slots$TimeSlot, 
                                levels = c("[6,9)", "[9,12)", 
                                           "[12,14)", "[14,17)",
                                           "[17,20)", "[20,22)", "[22, 6)"))

FoodbyCB_7slots <- Energy_slots %>% 
  group_by(ID, TimeSlot) %>% 
  summarise(sumEnergy = sum(Tot_Energ),
            sumCarb = sum(Tot_Carb), 
            sumSugar = sum(Tot_Sugar), 
            sumStarch = sum(Tot_Starch),
            sumFibre = sum(Tot_Fibre), 
            sumNMES = sum(Tot_NMES),
            sumFat = sum(Tot_Fat),
            sumProt = sum(Tot_Prot), 
            sumAlc = sum(Tot_Alc))

# Calculate the sum of each nutrients for each time slots
Carbsum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumCarb) %>% 
  spread(key = TimeSlot, 
         value = sumCarb)

Carbsum[is.na(Carbsum)] <- 0

names(Carbsum) <- c("ID", "Carb6_9", "Carb9_12", "Carb12_14", "Carb14_17",
                     "Carb17_20", "Carb20_22", "Carb22_6")

Energysum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumEnergy) %>% 
  spread(key = TimeSlot, 
         value = sumEnergy)
Energysum[is.na(Energysum)] <- 0

names(Energysum) <- c("ID", "Energy6_9", "Energy9_12", "Energy12_14", "Energy14_17",
                     "Energy17_20", "Energy20_22", "Energy22_6")

Starchsum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumStarch) %>% 
  spread(key = TimeSlot, 
         value = sumStarch)
Starchsum[is.na(Starchsum)] <- 0

names(Starchsum) <- c("ID", "Starch6_9", "Starch9_12", "Starch12_14", "Starch14_17",
                       "Starch17_20", "Starch20_22", "Starch22_6")

Sugarsum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumSugar) %>% 
  spread(key = TimeSlot, 
         value = sumSugar)
Sugarsum[is.na(Sugarsum)] <- 0

names(Sugarsum) <- c("ID", "Sugar6_9", "Sugar9_12", "Sugar12_14", "Sugar14_17",
                      "Sugar17_20", "Sugar20_22", "Sugar22_6")

Fibresum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumFibre) %>% 
  spread(key = TimeSlot, 
         value = sumFibre)
Fibresum[is.na(Fibresum)] <- 0

names(Fibresum) <- c("ID", "Fibre6_9", "Fibre9_12", "Fibre12_14", "Fibre14_17",
                      "Fibre17_20", "Fibre20_22", "Fibre22_6")

NMESsum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumNMES) %>% 
  spread(key = TimeSlot, 
         value = sumNMES)
NMESsum[is.na(NMESsum)] <- 0

names(NMESsum) <- c("ID", "NMES6_9", "NMES9_12", "NMES12_14", "NMES14_17",
                      "NMES17_20", "NMES20_22", "NMES22_6")

Fatsum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumFat) %>% 
  spread(key = TimeSlot, 
         value = sumFat)
Fatsum[is.na(Fatsum)] <- 0

names(Fatsum) <- c("ID", "Fat6_9", "Fat9_12", "Fat12_14", "Fat14_17",
                     "Fat17_20", "Fat20_22", "Fat22_6")

Protsum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumProt) %>% 
  spread(key = TimeSlot, 
         value = sumProt)
Protsum[is.na(Protsum)] <- 0

names(Protsum) <- c("ID", "Prot6_9", "Prot9_12", "Prot12_14", "Prot14_17",
                    "Prot17_20", "Prot20_22", "Prot22_6")

Alcsum <- FoodbyCB_7slots %>% 
  select(ID, TimeSlot, sumAlc) %>% 
  spread(key = TimeSlot, 
         value = sumAlc)
Alcsum[is.na(Alcsum)] <- 0

names(Alcsum) <- c("ID", "Alc6_9", "Alc9_12", "Alc12_14", "Alc14_17",
                     "Alc17_20", "Alc20_22", "Alc22_6")

# Extract number of days of diary completed

blood14 <- blood14 %>% 
  select(ID, Ndays)

blood56 <- blood56 %>% 
  select(ID, Ndays)

blood78 <- blood78 %>% 
  select(ID, NDays) %>% 
  rename(Ndays = NDays)

NDAYS <- rbind(blood14, blood56, blood78)
NDAYS$ID <- as.numeric(NDAYS$ID)

IntakeSlots <- Energysum %>% 
  left_join(Carbsum, by = "ID") %>% 
  left_join(Sugarsum, by = "ID") %>% 
  left_join(Starchsum, by = "ID") %>% 
  left_join(Fibresum, by = "ID") %>% 
  left_join(Fatsum, by = "ID") %>% 
  left_join(Protsum, by = "ID") %>% 
  left_join(NMESsum, by = "ID") %>% 
  left_join(Alcsum, by = "ID") %>% 
  left_join(NDAYS, by = "ID")

IntakeSlots$ID <- as.numeric(IntakeSlots$ID)

IntakeSlots$Energy6_9 <- IntakeSlots$Energy6_9/(IntakeSlots$Ndays)

for (i in 3:57){
  IntakeSlots[, i] <- IntakeSlots[, i]/(IntakeSlots$Ndays)
}

# select the variables needed, recoding 
# NAs, renaming to the same ----------------------------------------

BMI78 <- blood78 %>% 
  select(ID, Sex, age, bmival, wstval, Diabetes, bpmedc2, bpmedd2, 
         hyper140_2, hibp140_2,  Glucose, A1C, cigsta3, dnoft3, 
         dnnow, wti_Y78, wtn_Y78, wtb_Y78, ethgrp5, ethgrp2,
         cluster1, cluster2, cluster3, nssec8, paidemployment, 
         qual7, eqvinc, MarSt2, cluster4, cluster5, area, gor, 
         LDL, HDL, Chol, Trig) %>% 
  rename(wti = wti_Y78, wtn = wtn_Y78, wtb = wtb_Y78, 
         drink = dnoft3) %>% 
  mutate(Years = "7-8", MVPAtime = NA, MarStat = NA) %>% 
  replace_with_na(replace = list(bmival = -1, qual7 = -8,
                                 wstval = -1, eqvinc = -1,
                                 bpmedd2 = -1, MarSt2 = -1,
                                 bpmedc2 = -1, hyper140_2 = -7, 
                                 hibp140_2 = -7, Glucose = -1, 
                                 A1C  = -1,  LDL = -1, HDL = -1, 
                                 Chol = -1, Trig = -1, dnnow = -1,
                                 drink = -1, ethgrp5 = -4, 
                                 ethgrp2 = -4, cigsta3 = -1, 
                                 nssec8 = -9,  paidemployment = -1)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, 
                                 nssec8 = 99, qual7 = -1, drink = -8, 
                                 ethgrp2 = -9, ethgrp5 = -9)) %>% 
  replace_with_na(replace = list(drink = -9, ethgrp2 = -8, ethgrp5 = -8,
                                 cigsta3 = -8)) %>% 
  replace_with_na(replace = list(ethgrp2 = -1, ethgrp5 = -1))

BMI56 <- blood56 %>% 
  select(ID, Sex, age, area, bmival, wstval, Diabetes, bpmedc2, bpmedd2, 
         hyper140_2, hibp140_2, MVPAtime, Glucose, A1C, cigsta3, dnoft3, 
         dnnow, wti_Y56, wtn_Y56, wtb_Y56, ethgrp5, ethgrp2, qual7,
         cluster1, cluster2, cluster3, nssec8, paidemployment, eqvinc, 
         MarSt2, cluster4, cluster5, area, gor, LDL, HDL, Chol, Trig) %>% 
  mutate(Years = "5-6", MarStat = NA) %>% 
  rename(wti = wti_Y56, wtn = wtn_Y56, wtb = wtb_Y56, drink = dnoft3) %>% 
  replace_with_na(replace = list(bmival = -1, paidemployment = -9,
                                 wstval = -1, qual7 = -8,
                                 bpmedd2 = -1, eqvinc = -1,
                                 bpmedc2 = -1, MarSt2 = -1,
                                 hyper140_2 = -7, hibp140_2 = -7, 
                                 Glucose = -1,  A1C  = -1, 
                                 dnnow = -1, drink = -1, ethgrp5 = -4,
                                 ethgrp2 = -4, LDL = -1, HDL = -1, 
                                 Chol = -1, Trig = -1, cigsta3 = -1,
                                 MVPAtime = -1, nssec8 = -9)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, 
                                 paidemployment = -8, 
                                 drink = -8, nssec8 = 99)) %>% 
    replace_with_na(replace = list(drink = -9, paidemployment = -1, 
                                   qual7 = -1, cigsta3 = -8))

BMI14 <- blood14 %>% 
  select(ID, Sex, age, bmival, wstval, Diabetes, bpmedc, bpmedd, hyper140, 
         hibp140, MVPAtime, Glucose, A1C, cigsta3, dnoft3, dnnow, 
         wti_CY1234, wtn_CY1234, wtb_CY1234, ethgr5, ethgr2, cluster, 
         area, gor, nssec8, paidemployment, qual7, eqvinc, MarSt2, 
         MarStat, LDL, HDL, Chol, Trig) %>%
  rename(hyper140_2 = hyper140, hibp140_2 = hibp140, bpmedd2 = bpmedd, 
         bpmedc2 = bpmedc, cluster1 = cluster, ethgrp5 = ethgr5, 
         ethgrp2 = ethgr2, wti = wti_CY1234, wtn = wtn_CY1234, 
         wtb =  wtb_CY1234, drink = dnoft3) %>% 
  mutate(cluster2 = NA, cluster3 = NA, cluster4 = NA, cluster5 = NA, 
         Years = "1-4") %>% 
  replace_with_na(replace = list(bmival = -1, paidemployment = -9, 
                                 wstval = -1, qual7 = -8,
                                 bpmedd2 = -1, eqvinc = -1,
                                 bpmedc2 = -1, MarSt2 = -4, 
                                 hyper140_2 = -7, MarStat = -4,
                                 hibp140_2 = -7,  Glucose = -1, 
                                 A1C  = -1,  dnnow = -1, LDL = -1, 
                                 HDL = -1, Chol = -1, Trig = -1,
                                 drink = -1, ethgrp5 = -4, ethgrp2 = -4,
                                 cigsta3 = -1, MVPAtime = -4,  
                                 nssec8 = -8)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, 
                                 MVPAtime = -1, paidemployment = -8,
                                 drink = -8, nssec8 = -1, qual7 = -1, 
                                 MarSt2 = -1)) %>% 
  replace_with_na(replace = list(drink = -9, paidemployment = -4,
                                 cigsta3 = -8, nssec8 = 99)) %>% 
  replace_with_na(replace = list(paidemployment = -1))

BMI <- bind_rows(BMI14, BMI56, BMI78)

Energy14 <- food14 %>% 
  select(ID, Country, SurveyYear, EnergykJ, Carbohydrateg, 
         CHOpctotE, Proteing, ProteinpctotE, Alcoholg, AlcoholpctotE,
         Fatg, FatpctotE)
Energy14$Country[Energy14$Country == "Northern Ireland"] <- "NI"

Energy56 <- food56 %>% 
  select(ID, Country, Surveyyear, EnergykJ, Carbohydrateg, 
         CHOpctotE, Proteing, ProteinpctotE, Alcoholg, AlcoholpctotE,
         Fatg, FatpctotE) %>% 
  rename(SurveyYear = Surveyyear)

Energy78 <- food78 %>% 
  select(ID, Country, SurveyYear, EnergykJ, Carbohydrateg, 
         CHOpctotE, Proteing, ProteinpctotE, Alcoholg, AlcoholpctotE,
         Fatg, FatpctotE)

Energy <- bind_rows(Energy14, Energy56, Energy78)

dta_NDNS <- read_csv("dta_NDNS_Tslots.csv") # extract the day of week data

CW3idday$DayNo <- ave(CW3idday$ID_DAY, CW3idday$ID, 
                      FUN = seq_along) # adding the day no

dta_DayofWeek <- dta_NDNS %>% 
  select(id, id_dy, DayofWeek)

dta_DayofWeek$DayNo <- unlist(strsplit(dta_DayofWeek$id_dy,
                                       "D"))[c(FALSE, TRUE)] # creating a day No
names(dta_DayofWeek)[1] <- "ID"
dta_DayofWeek$DayNo <- as.numeric(dta_DayofWeek$DayNo)

CW3idday <-   CW3idday %>%  
  left_join(dta_DayofWeek, by= c("ID", "DayNo"))
